Estimating Neighborhood Rents with Scraped Data

### 2018-07-04 manuscript/index.Rmd
implemented use of load_data() helper throughout, all db queries now follow same structure via fn

Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist. The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

![Scatter plot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties. Purple dashed line indicates loess fit.](../output/rent_scatter_across_1B.png)

@@

>>>>

Figure 5 shows a bubbleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e. about 2/3 of the 124 tracts in Seattle using 2000 definitions). The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

![Bubbleplot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties. Purple dashed line indicates loess fit.](../output/rent_scatter_across_1B.png)

### 2018-07-04 manuscript/manuscript.Rmd
implemented use of load_data() helper throughout, all db queries now follow same structure via fn

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
date: "June 5, 2018"
output:  pdf_document
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for 2017 Q2 through 2018 Q1, while the test data are 2018 Q2 median values as observed through June 4th. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

@@

@@

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size. Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for 2017 Q2 upward in favor of regularly posted commercial listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75. Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between

@@

Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 75 tracts where comparable estimates exist. The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and good agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

To the extent that these smaller property units may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone. Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.

@@

The following models learn from Craigslist data accumulated up till the prior quarter (i.e. `r as.yearqtr(Sys.Date())-1`) to accomplish a few different goals. The most prominent goal is to use the panel data's temporal and spatial structure to understand variation in rent levels over time and across space. A secondary goal is to forecast the Seattle neighborhood rent surface for the current quarter. Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts.

@@

#table of fit statistics
fit_stats <- sea_df %>%
  mutate(train_test = ifelse(listingQtr == as.yearqtr(Sys.Date()), "Test", "Training")) %>%
  mutate(int_err = actualRent - int_Med,
         ns_err = actualRent - ns_Med,
         nsar1_err = actualRent - nsar1_Med,
         bym_err = actualRent - bym_Med,
         spt_err = actualRent - spt_Med) %>% 
  group_by(train_test) %>%
  summarize(int_rmse = rmse(int_err[!is.na(int_err)]),
            int_mae = mae(int_err[!is.na(int_err)]),
            ns_rmse = rmse(ns_err[!is.na(ns_err)]),
            ns_mae = mae(ns_err[!is.na(ns_err)]),
            nsar1_rmse = rmse(nsar1_err[!is.na(nsar1_err)]),
            nsar1_mae = mae(nsar1_err[!is.na(nsar1_err)]),
            bym_rmse = rmse(bym_err[!is.na(bym_err)]),
            bym_mae = mae(bym_err[!is.na(bym_err)]),
            spt_rmse = rmse(spt_err[!is.na(spt_err)]),
            spt_mae = mae(spt_err[!is.na(spt_err)])) %>%
  dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
```

The following tables show in-sample and out-of-sample (i.e. 2018 Q2) accuracy for the 5 model specifications in terms of root mean squared-error (RMSE) and mean absolute error (MAE). Note how though the Non-Spatial models (i.e. `ns` and `nsar1`) achieve slightly better in-sample fits, both Spatial models are more accurate at predicting the current month's surface of rent values. It is important to note how the goal in the present case is to understand the data generating process, that is, the systematic variation in the data less the random component. So while we want to minimize model error as a way to judge if our models improve on each other, we also want to smooth over sampling variability in the Craigslist data that induces local deviations compared to the true median rent. The [information criteria for the models](http://hesscl.com/smooth-sea/dashboard/index.html#model-fit) generally show that the spatial specifications are more likely models of the neighborhood rent data generating process.

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
output:  pdf_document
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

@@

Only the Craigslist and Dupre+Scott can offer insight into more recent median levels for the Seattle rental housing market. Both data suggest continued increases in the market-wide median rent for the first half of 2017, though the Craigslist slope is about twice in its size. Some of greater variation observed between the first and second data points of the Craigslist series may be related to missing listings (whether due to the early-stage process or outages), something which might bias the estimate for Q2 2017 upward in favor of regularly posted commercial listings. Though the estimates here reflect deduplicated listings, regularly updating a listing raises its likelihood of being at the top of the index our scraper works off of, thereby increasing its selection probability. This volatility in Craigslist asking rents could also be brought on by recently-developed buildings in "lease-up" adding a glut of expensive listings to the quarters' samples.

However, by the final point of Dupre+Scott's series (Fall 2017, corresponding to September), the difference compared to the Craigslist median was about $75. Mirroring other accounts about a decline in rents during Winter 2018, the Craigslist data showed a fairly precipitous decline between Q4 2017 and Q1 2018.

@@

@@

Figure 5 shows a scatter plot/bubble plot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e. about 2/3 of the 124 2000-definition tracts in Seattle). The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

To the extent that units in smaller properties (in terms of the number of units) may tend to be newer, larger in housing unit square footage or offer added amenities, the true value of rent for the neighborhood will be higher than what it is estimated at using 20+ unit properties alone. Similarly, the more that smaller properties account for the overall population of rental housing units, the more likely that higher values indicated in Craigslist are not erroneous but instead reflective of a broader set of residential units.

@@

@@

The following models learn from Craigslist data accumulated from Q1 2017 Q1 through Q1 2018 to accomplish a few different goals. The most prominent goal is to use the panel data's temporal and spatial structure to understand variation in rent levels over time and across space. A secondary goal is to forecast the Seattle neighborhood rent surface for the current quarter. Finally, the following models allow for relatively straightforward imputation of missing values based on time period and the rent values of proximate tracts.

@@

#table of fit statistics
fit_stats <- sea_df %>%
  mutate(train_test = ifelse(listingQtr == "2018 Q2", "Test", "Training")) %>%
  mutate(int_err = actualRent - int_Med,
         ns_err = actualRent - ns_Med,
         nsar1_err = actualRent - nsar1_Med,
         bym_err = actualRent - bym_Med,
         spt_err = actualRent - spt_Med) %>% 
  group_by(train_test) %>%
  summarize(int_rmse = rmse(int_err[!is.na(int_err)]),
            int_mae = mae(int_err[!is.na(int_err)]),
            ns_rmse = rmse(ns_err[!is.na(ns_err)]),
            ns_mae = mae(ns_err[!is.na(ns_err)]),
            nsar1_rmse = rmse(nsar1_err[!is.na(nsar1_err)]),
            nsar1_mae = mae(nsar1_err[!is.na(nsar1_err)]),
            bym_rmse = rmse(bym_err[!is.na(bym_err)]),
            bym_mae = mae(bym_err[!is.na(bym_err)]),
            spt_rmse = rmse(spt_err[!is.na(spt_err)]),
            spt_mae = mae(spt_err[!is.na(spt_err)])) %>%
  dplyr::select(train_test, ends_with("_rmse"), ends_with("_mae"))
```

The following tables show in-sample and out-of-sample (i.e. Q2 2018) accuracy for the 5 model specifications in terms of root mean squared-error (RMSE) and mean absolute error (MAE). Note how though the Non-Spatial models (i.e. `ns` and `nsar1`) achieve slightly better in-sample fits, both Spatial models are more accurate at predicting the current month's surface of rent values. It is important to note how the goal in the present case is to understand the data generating process, that is, the systematic variation in the data less the random component. So while we want to minimize model error as a way to judge if our models improve on each other, we also want to smooth over sampling variability in the Craigslist data that induces local deviations compared to the true median rent. The [information criteria for the models](http://hesscl.com/smooth-sea/dashboard/index.html#model-fit) generally show that the spatial specifications are more likely models of the neighborhood rent data generating process.
